---
title: "dry mass conversion"
author: "Liz"
date: "May 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Set up

Load necessary packages:

```{r,warning=FALSE}
library(readr)
library(dplyr)
library(lubridate)
```




Loading dataset:

```{r,message=FALSE}

para <- read_csv("~/Manduca expts/Summer+Fall 2016/CxPxT/data files/raw cpt.csv")
View(para)
```



Make timepoint data (3rd, 4th, 5th, emerged, etc), bin, and instars factors (others?)

Ex:

```{r}
para$timepoint.died=as.factor(para$timepoint.died)
para$timepoint.4=as.factor(para$timepoint.4)
```


```{r,include=FALSE}
para$timepoint.5=as.factor(para$timepoint.5)
para$timepoint.wander=as.factor(para$timepoint.wander)
para$timepoint.em=as.factor(para$timepoint.em)
para$bin=as.factor(para$bin)
para$instar.em=as.factor(para$instar.em)
para$instar.T0=as.factor(para$instar.T0)
para$instar.T1=as.factor(para$instar.T1)
para$instar.T2=as.factor(para$instar.T2)
para$instar.T3=as.factor(para$instar.T3)
para$instar.T4=as.factor(para$instar.T4)
para$instar.T5=as.factor(para$instar.T5)
para$instar.T6=as.factor(para$instar.T6)
para$instar.T7=as.factor(para$instar.T7)
para$instar.T8=as.factor(para$instar.T8)
para$instar.T9=as.factor(para$instar.T9)
para$instar.T10=as.factor(para$instar.T10)
para$instar.T11=as.factor(para$instar.T11)
para$instar.T12=as.factor(para$instar.T12)
para$instar.T13=as.factor(para$instar.T13)
para$instar.T14=as.factor(para$instar.T14)
para$instar.T15=as.factor(para$instar.T15)
para$instar.T16=as.factor(para$instar.T16)
para$instar.T17=as.factor(para$instar.T17)
para$instar.T18=as.factor(para$instar.T18)
para$instar.T19=as.factor(para$instar.T19)
para$instar.T20=as.factor(para$instar.T20)
para$instar.T21=as.factor(para$instar.T21)
para$instar.T22=as.factor(para$instar.T22)
para$instar.T23=as.factor(para$instar.T23)
para$instar.T24=as.factor(para$instar.T24)
para$instar.T25=as.factor(para$instar.T25)
para$instar.T26=as.factor(para$instar.T26)
para$instar.T27=as.factor(para$instar.T27)
para$instar.T28=as.factor(para$instar.T28)
para$instar.T29=as.factor(para$instar.T29)
para$instar.T30=as.factor(para$instar.T30)
para$instar.T31=as.factor(para$instar.T31)
para$instar.T32=as.factor(para$instar.T32)
para$instar.T33=as.factor(para$instar.T33)
para$instar.T34=as.factor(para$instar.T34)
para$instar.T35=as.factor(para$instar.T35)
para$instar.T36=as.factor(para$instar.T36)
para$instar.T37=as.factor(para$instar.T37)
para$instar.T38=as.factor(para$instar.T38)
```



###Converting initial wet masses to dry mass


Create new columns with initial caterpillar diet wet mass * the slope of the initial regression line

```{r,include=FALSE}
para$dry.diet.cat.in.T0<-NULL
para$dry.diet.cat.out.T0<-NULL
para$dry.diet.cont.inT0<-NULL
para$dry.diet.cont.out.T0<-NULL
para$comments<-NULL
```


Ex: 
```{r}
para$dry.cat.in.T0<-(para$diet.cat.in.T0*.227)
para$dry.cat.in.T1<-(para$diet.cat.in.T1*.227)
```

```{r,include=FALSE}
para$dry.cat.in.T2<-(para$diet.cat.in.T2*.227)
para$dry.cat.in.T3<-(para$diet.cat.in.T3*.227)
para$dry.cat.in.T4<-(para$diet.cat.in.T4*.227)
para$dry.cat.in.T5<-(para$diet.cat.in.T5*.227)
para$dry.cat.in.T6<-(para$diet.cat.in.T6*.227)
para$dry.cat.in.T7<-(para$diet.cat.in.T7*.227)
para$dry.cat.in.T8<-(para$diet.cat.in.T8*.227)
para$dry.cat.in.T9<-(para$diet.cat.in.T9*.227)
para$dry.cat.in.T10<-(para$diet.cat.in.T10*.227)
para$dry.cat.in.T11<-(para$diet.cat.in.T11*.227)
para$dry.cat.in.T12<-(para$diet.cat.in.T12*.227)
para$dry.cat.in.T13<-(para$diet.cat.in.T13*.227)
para$dry.cat.in.T14<-(para$diet.cat.in.T14*.227)
para$dry.cat.in.T15<-(para$diet.cat.in.T15*.227)
para$dry.cat.in.T16<-(para$diet.cat.in.T16*.227)
para$dry.cat.in.T17<-(para$diet.cat.in.T17*.227)
para$dry.cat.in.T18<-(para$diet.cat.in.T18*.227)
para$dry.cat.in.T19<-(para$diet.cat.in.T19*.227)
para$dry.cat.in.T20<-(para$diet.cat.in.T20*.227)
para$dry.cat.in.T21<-(para$diet.cat.in.T21*.227)
para$dry.cat.in.T22<-(para$diet.cat.in.T22*.227)
para$dry.cat.in.T23<-(para$diet.cat.in.T23*.227)
para$dry.cat.in.T24<-(para$diet.cat.in.T24*.227)
para$dry.cat.in.T25<-(para$diet.cat.in.T25*.227)
para$dry.cat.in.T26<-(para$diet.cat.in.T26*.227)
para$dry.cat.in.T27<-(para$diet.cat.in.T27*.227)
para$dry.cat.in.T28<-(para$diet.cat.in.T28*.227)
para$dry.cat.in.T29<-(para$diet.cat.in.T29*.227)
para$dry.cat.in.T30<-(para$diet.cat.in.T30*.227)
para$dry.cat.in.T31<-(para$diet.cat.in.T31*.227)
para$dry.cat.in.T32<-(para$diet.cat.in.T32*.227)
para$dry.cat.in.T33<-(para$diet.cat.in.T33*.227)
para$dry.cat.in.T34<-(para$diet.cat.in.T34*.227)
para$dry.cat.in.T35<-(para$diet.cat.in.T35*.227)
para$dry.cat.in.T36<-(para$diet.cat.in.T36*.227)
para$dry.cat.in.T37<-(para$diet.cat.in.T37*.227)

```




Create new columns with initial control diet wet mass * the slope of the initial regression line

Ex: 
```{r}
para$dry.cont.in.T0<-(para$diet.cont.in.T0*.227)
para$dry.cont.in.T1<-(para$diet.cont.in.T1*.227)
```


```{r,include=FALSE}
para$dry.cont.in.T2<-(para$diet.cont.in.T2*.227)
para$dry.cont.in.T3<-(para$diet.cont.in.T3*.227)
para$dry.cont.in.T4<-(para$diet.cont.in.T4*.227)
para$dry.cont.in.T5<-(para$diet.cont.in.T5*.227)
para$dry.cont.in.T6<-(para$diet.cont.in.T6*.227)
para$dry.cont.in.T7<-(para$diet.cont.in.T7*.227)
para$dry.cont.in.T8<-(para$diet.cont.in.T8*.227)
para$dry.cont.in.T9<-(para$diet.cont.in.T9*.227)
para$dry.cont.in.T10<-(para$diet.cont.in.T10*.227)
para$dry.cont.in.T11<-(para$diet.cont.in.T11*.227)
para$dry.cont.in.T12<-(para$diet.cont.in.T12*.227)
para$dry.cont.in.T13<-(para$diet.cont.in.T13*.227)
para$dry.cont.in.T14<-(para$diet.cont.in.T14*.227)
para$dry.cont.in.T15<-(para$diet.cont.in.T15*.227)
para$dry.cont.in.T16<-(para$diet.cont.in.T16*.227)
para$dry.cont.in.T17<-(para$diet.cont.in.T17*.227)
para$dry.cont.in.T18<-(para$diet.cont.in.T18*.227)
para$dry.cont.in.T19<-(para$diet.cont.in.T19*.227)
para$dry.cont.in.T20<-(para$diet.cont.in.T20*.227)
para$dry.cont.in.T21<-(para$diet.cont.in.T21*.227)
para$dry.cont.in.T22<-(para$diet.cont.in.T22*.227)
para$dry.cont.in.T23<-(para$diet.cont.in.T23*.227)
para$dry.cont.in.T24<-(para$diet.cont.in.T24*.227)
para$dry.cont.in.T25<-(para$diet.cont.in.T25*.227)
para$dry.cont.in.T26<-(para$diet.cont.in.T26*.227)
para$dry.cont.in.T27<-(para$diet.cont.in.T27*.227)
para$dry.cont.in.T28<-(para$diet.cont.in.T28*.227)
para$dry.cont.in.T29<-(para$diet.cont.in.T29*.227)
para$dry.cont.in.T30<-(para$diet.cont.in.T30*.227)
para$dry.cont.in.T31<-(para$diet.cont.in.T31*.227)
para$dry.cont.in.T32<-(para$diet.cont.in.T32*.227)
para$dry.cont.in.T33<-(para$diet.cont.in.T33*.227)
para$dry.cont.in.T34<-(para$diet.cont.in.T34*.227)
para$dry.cont.in.T35<-(para$diet.cont.in.T35*.227)
para$dry.cont.in.T36<-(para$diet.cont.in.T36*.227)
para$dry.cont.in.T37<-(para$diet.cont.in.T37*.227)

```







###Converting final wet masses to dry masses



Create new columns with final caterpillar diet wet mass * the slope of the final regression line


Ex: 
```{r}
para$dry.cat.out.T0<-(para$diet.cat.out.T0*.237)
para$dry.cat.out.T1<-(para$diet.cat.out.T1*.237)
```



```{r,include=FALSE}
para$dry.cat.out.T2<-(para$diet.cat.out.T2*.237)
para$dry.cat.out.T3<-(para$diet.cat.out.T3*.237)
para$dry.cat.out.T4<-(para$diet.cat.out.T4*.237)
para$dry.cat.out.T5<-(para$diet.cat.out.T5*.237)
para$dry.cat.out.T6<-(para$diet.cat.out.T6*.237)
para$dry.cat.out.T7<-(para$diet.cat.out.T7*.237)
para$dry.cat.out.T8<-(para$diet.cat.out.T8*.237)
para$dry.cat.out.T9<-(para$diet.cat.out.T9*.237)
para$dry.cat.out.T10<-(para$diet.cat.out.T10*.237)
para$dry.cat.out.T11<-(para$diet.cat.out.T11*.237)
para$dry.cat.out.T12<-(para$diet.cat.out.T12*.237)
para$dry.cat.out.T13<-(para$diet.cat.out.T13*.237)
para$dry.cat.out.T14<-(para$diet.cat.out.T14*.237)
para$dry.cat.out.T15<-(para$diet.cat.out.T15*.237)
para$dry.cat.out.T16<-(para$diet.cat.out.T16*.237)
para$dry.cat.out.T17<-(para$diet.cat.out.T17*.237)
para$dry.cat.out.T18<-(para$diet.cat.out.T18*.237)
para$dry.cat.out.T19<-(para$diet.cat.out.T19*.237)
para$dry.cat.out.T20<-(para$diet.cat.out.T20*.237)
para$dry.cat.out.T21<-(para$diet.cat.out.T21*.237)
para$dry.cat.out.T22<-(para$diet.cat.out.T22*.237)
para$dry.cat.out.T23<-(para$diet.cat.out.T23*.237)
para$dry.cat.out.T24<-(para$diet.cat.out.T24*.237)
para$dry.cat.out.T25<-(para$diet.cat.out.T25*.237)
para$dry.cat.out.T26<-(para$diet.cat.out.T26*.237)
para$dry.cat.out.T27<-(para$diet.cat.out.T27*.237)
para$dry.cat.out.T28<-(para$diet.cat.out.T28*.237)
para$dry.cat.out.T29<-(para$diet.cat.out.T29*.237)
para$dry.cat.out.T30<-(para$diet.cat.out.T30*.237)
para$dry.cat.out.T31<-(para$diet.cat.out.T31*.237)
para$dry.cat.out.T32<-(para$diet.cat.out.T32*.237)
para$dry.cat.out.T33<-(para$diet.cat.out.T33*.237)
para$dry.cat.out.T34<-(para$diet.cat.out.T34*.237)
para$dry.cat.out.T35<-(para$diet.cat.out.T35*.237)
para$dry.cat.out.T36<-(para$diet.cat.out.T36*.237)

```



Create new columns with final control diet wet mass * the slope of the final regression line


Ex: 
```{r}
para$dry.cont.out.T0<-(para$diet.cont.out.T0*.237)
para$dry.cont.out.T1<-(para$diet.cont.out.T1*.237)
```



```{r,include=FALSE}
para$dry.cont.out.T2<-(para$diet.cont.out.T2*.237)
para$dry.cont.out.T3<-(para$diet.cont.out.T3*.237)
para$dry.cont.out.T4<-(para$diet.cont.out.T4*.237)
para$dry.cont.out.T5<-(para$diet.cont.out.T5*.237)
para$dry.cont.out.T6<-(para$diet.cont.out.T6*.237)
para$dry.cont.out.T7<-(para$diet.cont.out.T7*.237)
para$dry.cont.out.T8<-(para$diet.cont.out.T8*.237)
para$dry.cont.out.T9<-(para$diet.cont.out.T9*.237)
para$dry.cont.out.T10<-(para$diet.cont.out.T10*.237)
para$dry.cont.out.T11<-(para$diet.cont.out.T11*.237)
para$dry.cont.out.T12<-(para$diet.cont.out.T12*.237)
para$dry.cont.out.T13<-(para$diet.cont.out.T13*.237)
para$dry.cont.out.T14<-(para$diet.cont.out.T14*.237)
para$dry.cont.out.T15<-(para$diet.cont.out.T15*.237)
para$dry.cont.out.T16<-(para$diet.cont.out.T16*.237)
para$dry.cont.out.T17<-(para$diet.cont.out.T17*.237)
para$dry.cont.out.T18<-(para$diet.cont.out.T18*.237)
para$dry.cont.out.T19<-(para$diet.cont.out.T19*.237)
para$dry.cont.out.T20<-(para$diet.cont.out.T20*.237)
para$dry.cont.out.T21<-(para$diet.cont.out.T21*.237)
para$dry.cont.out.T22<-(para$diet.cont.out.T22*.237)
para$dry.cont.out.T23<-(para$diet.cont.out.T23*.237)
para$dry.cont.out.T24<-(para$diet.cont.out.T24*.237)
para$dry.cont.out.T25<-(para$diet.cont.out.T25*.237)
para$dry.cont.out.T26<-(para$diet.cont.out.T26*.237)
para$dry.cont.out.T27<-(para$diet.cont.out.T27*.237)
para$dry.cont.out.T28<-(para$diet.cont.out.T28*.237)
para$dry.cont.out.T29<-(para$diet.cont.out.T29*.237)
para$dry.cont.out.T30<-(para$diet.cont.out.T30*.237)
para$dry.cont.out.T31<-(para$diet.cont.out.T31*.237)
para$dry.cont.out.T32<-(para$diet.cont.out.T32*.237)
para$dry.cont.out.T33<-(para$diet.cont.out.T33*.237)
para$dry.cont.out.T34<-(para$diet.cont.out.T34*.237)
para$dry.cont.out.T35<-(para$diet.cont.out.T35*.237)
para$dry.cont.out.T36<-(para$diet.cont.out.T36*.237)

```





###Calculating stand alone and running total consumption for each time point/time period


Create a new column with consumption of first time period (cat. initial - cat. final: in.T0-out.T0)

```{r}
para$dry.cnsmp.T0_T1<-(para$dry.cat.in.T0-para$dry.cat.out.T0)
```



Create a new column with running total of consumption for the first time period--will be the same value as the stand alone consumption for the first time period

```{r}
para$dry.totcnsmp.T1<-(para$dry.cat.in.T0-para$dry.cat.out.T0)
```



Create a new column with stand alone consumption for the 2nd time period (in.T1-out.T1)

```{r}
para$dry.cnsmp.T1_T2<-(para$dry.cat.in.T1-para$dry.cat.out.T1)
```



Create a new column with the running total of consumption for 2nd time period--will be the runnng total from the 1st tp + the stand alone total from the 2nd tp (dry.totcnsmp.T1 + dry.cnsmp.T1-T2)  

```{r}
para$dry.totcnsmp.T2<-(para$dry.totcnsmp.T1+para$dry.cnsmp.T1_T2)
```



Create new columns for stand alone and running total consumption for all time points

```{r,include=FALSE}
#3rd time period stand alone (in.T2-out.T2)

para$dry.cnsmp.T2_T3<-(para$dry.cat.in.T2-para$dry.cat.out.T2)

#3rd time period running total (dry.totcnsmp.T2 + dry.cnsmp.T2_T3)

para$dry.totcnsmp.T3<-(para$dry.totcnsmp.T2+para$dry.cnsmp.T2_T3)

#4th time period stand alone (in.T3-out.T3)

para$dry.cnsmp.T3_T4<-(para$dry.cat.in.T3-para$dry.cat.out.T3)

#4th time period running total (dry.totcnsmp.T3 + dry.cnsmp.T3_T4)

para$dry.totcnsmp.T4<-(para$dry.totcnsmp.T3+para$dry.cnsmp.T3_T4)

#5th time period stand alone (in.T4-out.T4)

para$dry.cnsmp.T4_T5<-(para$dry.cat.in.T4-para$dry.cat.out.T4)

#5th time period running total (dry.totcnsmp.T4 + dry.cnsmp.T4_T5)

para$dry.totcnsmp.T5<-(para$dry.cnsmp.T4_T5+para$dry.totcnsmp.T4)

#6th time period stand alone (in.T5-out.T5)

para$dry.cnsmp.T5_T6<-(para$dry.cat.in.T5-para$dry.cat.out.T5)

#6th time period running total (dry.totcnsmp.T5 + dry.cnsmp.T5_T6)

para$dry.totcnsmp.T6<-(para$dry.cnsmp.T5_T6+para$dry.totcnsmp.T5)

#7th time period stand alone (in.T6-out.T6)

para$dry.cnsmp.T6_T7<-(para$dry.cat.in.T6-para$dry.cat.out.T6)

#7th time period running total (dry.cnsmp.T6_T7 + dry.tot.cnsmp.T6)

para$dry.totcnsmp.T7<-(para$dry.cnsmp.T6_T7+para$dry.totcnsmp.T6)

#8th time period stand alone (in.T7-out.T7)

para$dry.cnsmp.T7_T8<-(para$dry.cat.in.T7-para$dry.cat.out.T7)

#8th time period running total (dry.cnsmp.T7_T8 + dry.tot.cnsmp.T7)

para$dry.totcnsmp.T8<-(para$dry.cnsmp.T7_T8+para$dry.totcnsmp.T7)

#9th time period stand alone (in.T8-out.T8)

para$dry.cnsmp.T8_T9<-(para$dry.cat.in.T8-para$dry.cat.out.T8)

#9th time period running total (T8_T9 + tot.T8)

para$dry.totcnsmp.T9<-(para$dry.cnsmp.T8_T9+para$dry.totcnsmp.T8)

#10th time period stand alone (in.T9-out.T9)

para$dry.cnsmp.T9_T10<-(para$dry.cat.in.T9-para$dry.cat.out.T9)

#10th time period running total (T9_T10 + tot.T9)

para$dry.totcnsmp.T10<-(para$dry.cnsmp.T9_T10+para$dry.totcnsmp.T9)

#11th time period stand alone (in.T10-out.T10)

para$dry.cnsmp.T10_T11<-(para$dry.cat.in.T10-para$dry.cat.out.T10)

#11th time period running total (T10_T11 + tot.T10)

para$dry.totcnsmp.T11<-(para$dry.cnsmp.T10_T11+para$dry.totcnsmp.T10)

#12th time period stand alone (in.T11-out.T11)

para$dry.cnsmp.T11_T12<-(para$dry.cat.in.T11-para$dry.cat.out.T11)

#12th time period running total (T11_T12 + tot.T11)

para$dry.totcnsmp.T12<-(para$dry.cnsmp.T11_T12+para$dry.totcnsmp.T11)

#13th time period stand alone (in.T12-out.T12)

para$dry.cnsmp.T12_T13<-(para$dry.cat.in.T12-para$dry.cat.out.T12)

#13th time period running total (T12_T13 + tot.T12)

para$dry.totcnsmp.T13<-(para$dry.cnsmp.T12_T13+para$dry.totcnsmp.T12)

#14th time period stand alone (in.T13-out.T13)

para$dry.cnsmp.T13_T14<-(para$dry.cat.in.T13-para$dry.cat.out.T13)

#14th time period running total (T13_T14 + tot.T13)

para$dry.totcnsmp.T14<-(para$dry.cnsmp.T13_T14+para$dry.totcnsmp.T13)

#15th time period stand alone (in.T14-out.T14)

para$dry.cnsmp.T14_T15<-(para$dry.cat.in.T14-para$dry.cat.out.T14)

#15th time period running total (T14_T15 + tot.T14)

para$dry.totcnsmp.T15<-(para$dry.cnsmp.T14_T15+para$dry.totcnsmp.T14)

#16th time period stand alone (in.T15-out.T15)

para$dry.cnsmp.T15_T16<-(para$dry.cat.in.T15-para$dry.cat.out.T15)

#16th time period running total (T15_T16 + tot.T15)

para$dry.totcnsmp.T16<-(para$dry.cnsmp.T15_T16+para$dry.totcnsmp.T15)

#17th time period stand alone (in.T16-out.T16)

para$dry.cnsmp.T16_T17<-(para$dry.cat.in.T16-para$dry.cat.out.T16)

#17th time period running total (T16_T16 + tot.T16)

para$dry.totcnsmp.T17<-(para$dry.cnsmp.T16_T17+para$dry.totcnsmp.T16)

#18th time period stand alone (in.T17-out.T17)

para$dry.cnsmp.T17_T18<-(para$dry.cat.in.T17-para$dry.cat.out.T17)

#18th time period running total (T17_T18 + tot.T17)

para$dry.totcnsmp.T18<-(para$dry.cnsmp.T17_T18+para$dry.totcnsmp.T17)

#19th time period stand alone (in.T18-out.T18)

para$dry.cnsmp.T18_T19<-(para$dry.cat.in.T18-para$dry.cat.out.T18)

#19th time period running total (T18_T19 + tot.T18)

para$dry.totcnsmp.T19<-(para$dry.cnsmp.T18_T19+para$dry.totcnsmp.T18)

#20th time period stand alone (in.T19-out.T19)

para$dry.cnsmp.T19_T20<-(para$dry.cat.in.T19-para$dry.cat.out.T19)

#20th time period running total (T19_T20 + tot.T19)

para$dry.totcnsmp.T20<-(para$dry.cnsmp.T19_T20+para$dry.totcnsmp.T19)

#21st time period stand alone (in.T20-out.T20)

para$dry.cnsmp.T20_T21<-(para$dry.cat.in.T20-para$dry.cat.out.T20)

#21st time period running total (T20_T21 + tot.T20)

para$dry.totcnsmp.T21<-(para$dry.cnsmp.T20_T21+para$dry.totcnsmp.T20)

#22nd time period stand alone (in.T21-out.T21)

para$dry.cnsmp.T21_T22<-(para$dry.cat.in.T21-para$dry.cat.out.T21)

#22nd time period runnint total (T21_T22 + tot.T21)

para$dry.totcnsmp.T22<-(para$dry.cnsmp.T21_T22+para$dry.totcnsmp.T21)

#23rd time period stand alone (in.T22-out.T22)

para$dry.cnsmp.T22_T23<-(para$dry.cat.in.T22-para$dry.cat.out.T22)

#23rd time period running total (T22_T23 + tot.T22)

para$dry.totcnsmp.T23<-(para$dry.cnsmp.T22_T23+para$dry.totcnsmp.T22)

#24th time period stand alone (in.T23-out.T23)

para$dry.cnsmp.T23_T24<-(para$dry.cat.in.T23-para$dry.cat.out.T23)

#24th time period running total (T23_T24 + tot.T23)

para$dry.totcnsmp.T24<-(para$dry.cnsmp.T23_T24+para$dry.totcnsmp.T23)

#25th time period stand alone (in.T24-out.T24)

para$dry.cnsmp.T24_T25<-(para$dry.cat.in.T24-para$dry.cat.out.T24)

#25th time period running total (T24_T25 + tot.T24)

para$dry.totcnsmp.T25<-(para$dry.cnsmp.T24_T25+para$dry.totcnsmp.T24)

#26th time period stand alone (in.T25-out.T25)

para$dry.cnsmp.T25_T26<-(para$dry.cat.in.T25-para$dry.cat.out.T25)

#26th time period running total (T25_T26 + tot.T25)

para$dry.totcnsmp.T26<-(para$dry.cnsmp.T25_T26+para$dry.totcnsmp.T25)

#27th time period stand alone (in.T26-out.T26)

para$dry.cnsmp.T26_T27<-(para$dry.cat.in.T26-para$dry.cat.out.T26)

#27th time period running total (T26_T27 + tot.T26)

para$dry.totcnsmp.T27<-(para$dry.cnsmp.T26_T27+para$dry.totcnsmp.T26)

#28th time period stand alone (in.T27-out.T27)

para$dry.cnsmp.T27_T28<-(para$dry.cat.in.T27-para$dry.cat.out.T27)

#28th time period running total (T27_T28 + tot.T27)

para$dry.totcnsmp.T28<-(para$dry.cnsmp.T27_T28+para$dry.totcnsmp.T27)

#29th time period stand alone (in.T28-out.T28)

para$dry.cnsmp.T28_T29<-(para$dry.cat.in.T28-para$dry.cat.out.T28)

#29th time period running total (T28_T29 + tot.T28)

para$dry.totcnsmp.T29<-(para$dry.cnsmp.T28_T29+para$dry.totcnsmp.T28)

#30th time period stand alone (in.T29-out.T29)

para$dry.cnsmp.T29_T30<-(para$dry.cat.in.T29-para$dry.cat.out.T29)

#30th time period running total (T29_T30 + tot.T29)

para$dry.totcnsmp.T30<-(para$dry.cnsmp.T29_T30+para$dry.totcnsmp.T29)

#31st time period stand alone (in.T30-out.T30)

para$dry.cnsmp.T30_T31<-(para$dry.cat.in.T30-para$dry.cat.out.T30)

#31st time period running total (T30_T31 + tot.T30)

para$dry.totcnsmp.T31<-(para$dry.cnsmp.T30_T31+para$dry.totcnsmp.T30)

#32nd time period stand alone (in.T31-out.T31)

para$dry.cnsmp.T31_T32<-(para$dry.cat.in.T31-para$dry.cat.out.T31)

#32nd time period running total (T31_T32 + tot.T31)

para$dry.totcnsmp.T32<-(para$dry.cnsmp.T31_T32+para$dry.totcnsmp.T31)

#33rd time period stand alone (in.T32-out.T32)

para$dry.cnsmp.T32_T33<-(para$dry.cat.in.T32-para$dry.cat.out.T32)

#33rd time period running total (T32_T33 + tot.T32)

para$dry.totcnsmp.T33<-(para$dry.cnsmp.T32_T33+para$dry.totcnsmp.T32)

#34th time period stand alone (in.T33-out.T33)

para$dry.cnsmp.T33_T34<-(para$dry.cat.in.T33-para$dry.cat.out.T33)

#34th time period running total (T33_T34 + tot.T33)

para$dry.totcnsmp.T34<-(para$dry.cnsmp.T33_T34+para$dry.totcnsmp.T33)

#35th time period stand alone (in.T34-out.T34)

para$dry.cnsmp.T34_T35<-(para$dry.cat.in.T34-para$dry.cat.out.T34)

#35th time period running total (T34_T35 + tot.T34)

para$dry.totcnsmp.T35<-(para$dry.cnsmp.T34_T35+para$dry.totcnsmp.T34)

#36th time period stand alone (in.T35-out.T35)

para$dry.cnsmp.T35_T36<-(para$dry.cat.in.T35-para$dry.cat.out.T35)

#36th time period running total (T35_T36 + tot.T35)

para$dry.totcnsmp.T36<-(para$dry.cnsmp.T35_T36+para$dry.totcnsmp.T35)

#37th time period stand alone (in.T36-out.T36)

para$dry.cnsmp.T36_T37<-(para$dry.cat.in.T36-para$dry.cat.out.T36)

#37th time period running total (T36_T37 + tot.T36)

para$dry.totcnsmp.T37<-(para$dry.cnsmp.T36_T37+para$dry.totcnsmp.T36)


```



Making negative values = 1, so that models and graphs with log(tot.cnsmp) will run 

EX:

```{r}

para$dry.cnsmp.T0_T1[para$dry.cnsmp.T0_T1 < 0] <- 1
para$dry.cnsmp.T1_T2[para$dry.cnsmp.T1_T2 < 0] <- 1

para$dry.totcnsmp.T1[para$dry.totcnsmp.T1 < 0] <- 1
para$dry.totcnsmp.T2[para$dry.totcnsmp.T2 < 0] <- 1

```


```{r,include=FALSE}

para$dry.cnsmp.T2_T3[para$dry.cnsmp.T2_T3 < 0] <- 1
para$dry.cnsmp.T3_T4[para$dry.cnsmp.T3_T4 < 0] <- 1
para$dry.cnsmp.T4_T5[para$dry.cnsmp.T4_T5 < 0] <- 1
para$dry.cnsmp.T5_T6[para$dry.cnsmp.T5_T6 < 0] <- 1
para$dry.cnsmp.T6_T7[para$dry.cnsmp.T6_T7 < 0] <- 1
para$dry.cnsmp.T7_T8[para$dry.cnsmp.T7_T8 < 0] <- 1
para$dry.cnsmp.T8_T9[para$dry.cnsmp.T8_T9 < 0] <- 1
para$dry.cnsmp.T9_T10[para$dry.cnsmp.T9_T10 < 0] <- 1
para$dry.cnsmp.T10_T11[para$dry.cnsmp.T10_T11 < 0] <- 1
para$dry.cnsmp.T11_T12[para$dry.cnsmp.T11_T12 < 0] <- 1
para$dry.cnsmp.T12_T13[para$dry.cnsmp.T12_T13 < 0] <- 1
para$dry.cnsmp.T13_T14[para$dry.cnsmp.T13_T14 < 0] <- 1
para$dry.cnsmp.T14_T15[para$dry.cnsmp.T14_T15 < 0] <- 1
para$dry.cnsmp.T15_T16[para$dry.cnsmp.T15_T16 < 0] <- 1
para$dry.cnsmp.T16_T17[para$dry.cnsmp.T16_T17 < 0] <- 1
para$dry.cnsmp.T17_T18[para$dry.cnsmp.T17_T18 < 0] <- 1
para$dry.cnsmp.T18_T19[para$dry.cnsmp.T18_T19 < 0] <- 1
para$dry.cnsmp.T19_T20[para$dry.cnsmp.T19_T20 < 0] <- 1
para$dry.cnsmp.T20_T21[para$dry.cnsmp.T20_T21 < 0] <- 1
para$dry.cnsmp.T21_T22[para$dry.cnsmp.T21_T22 < 0] <- 1
para$dry.cnsmp.T22_T23[para$dry.cnsmp.T22_T23 < 0] <- 1
para$dry.cnsmp.T23_T24[para$dry.cnsmp.T23_T24 < 0] <- 1
para$dry.cnsmp.T24_T25[para$dry.cnsmp.T24_T25 < 0] <- 1
para$dry.cnsmp.T25_T26[para$dry.cnsmp.T25_T26 < 0] <- 1
para$dry.cnsmp.T26_T27[para$dry.cnsmp.T26_T27 < 0] <- 1
para$dry.cnsmp.T27_T28[para$dry.cnsmp.T27_T28 < 0] <- 1
para$dry.cnsmp.T28_T29[para$dry.cnsmp.T28_T29 < 0] <- 1
para$dry.cnsmp.T29_T30[para$dry.cnsmp.T29_T30 < 0] <- 1
para$dry.cnsmp.T30_T31[para$dry.cnsmp.T30_T31 < 0] <- 1
para$dry.cnsmp.T31_T32[para$dry.cnsmp.T31_T32 < 0] <- 1
para$dry.cnsmp.T32_T33[para$dry.cnsmp.T32_T33 < 0] <- 1
para$dry.cnsmp.T33_T34[para$dry.cnsmp.T33_T34 < 0] <- 1
para$dry.cnsmp.T34_T35[para$dry.cnsmp.T34_T35 < 0] <- 1
para$dry.cnsmp.T35_T36[para$dry.cnsmp.T35_T36 < 0] <- 1
para$dry.cnsmp.T36_T37[para$dry.cnsmp.T36_T37 < 0] <- 1




para$dry.totcnsmp.T3[para$dry.totcnsmp.T3 < 0] <- 1
para$dry.totcnsmp.T4[para$dry.totcnsmp.T4 < 0] <- 1
para$dry.totcnsmp.T5[para$dry.totcnsmp.T5 < 0] <- 1
para$dry.totcnsmp.T6[para$dry.totcnsmp.T6 < 0] <- 1
para$dry.totcnsmp.T7[para$dry.totcnsmp.T7 < 0] <- 1
para$dry.totcnsmp.T8[para$dry.totcnsmp.T8 < 0] <- 1
para$dry.totcnsmp.T9[para$dry.totcnsmp.T9 < 0] <- 1
para$dry.totcnsmp.T10[para$dry.totcnsmp.T10 < 0] <- 1
para$dry.totcnsmp.T11[para$dry.totcnsmp.T11 < 0] <- 1
para$dry.totcnsmp.T12[para$dry.totcnsmp.T12 < 0] <- 1
para$dry.totcnsmp.T13[para$dry.totcnsmp.T13 < 0] <- 1
para$dry.totcnsmp.T14[para$dry.totcnsmp.T14 < 0] <- 1
para$dry.totcnsmp.T15[para$dry.totcnsmp.T15 < 0] <- 1
para$dry.totcnsmp.T16[para$dry.totcnsmp.T16 < 0] <- 1
para$dry.totcnsmp.T17[para$dry.totcnsmp.T17 < 0] <- 1
para$dry.totcnsmp.T18[para$dry.totcnsmp.T18 < 0] <- 1
para$dry.totcnsmp.T19[para$dry.totcnsmp.T19 < 0] <- 1
para$dry.totcnsmp.T20[para$dry.totcnsmp.T20 < 0] <- 1
para$dry.totcnsmp.T21[para$dry.totcnsmp.T21 < 0] <- 1
para$dry.totcnsmp.T22[para$dry.totcnsmp.T22 < 0] <- 1
para$dry.totcnsmp.T23[para$dry.totcnsmp.T23 < 0] <- 1
para$dry.totcnsmp.T24[para$dry.totcnsmp.T24 < 0] <- 1
para$dry.totcnsmp.T25[para$dry.totcnsmp.T25 < 0] <- 1
para$dry.totcnsmp.T26[para$dry.totcnsmp.T26 < 0] <- 1
para$dry.totcnsmp.T27[para$dry.totcnsmp.T27 < 0] <- 1
para$dry.totcnsmp.T28[para$dry.totcnsmp.T28 < 0] <- 1
para$dry.totcnsmp.T29[para$dry.totcnsmp.T29 < 0] <- 1
para$dry.totcnsmp.T30[para$dry.totcnsmp.T30 < 0] <- 1
para$dry.totcnsmp.T31[para$dry.totcnsmp.T31 < 0] <- 1
para$dry.totcnsmp.T32[para$dry.totcnsmp.T32 < 0] <- 1
para$dry.totcnsmp.T33[para$dry.totcnsmp.T33 < 0] <- 1
para$dry.totcnsmp.T34[para$dry.totcnsmp.T34 < 0] <- 1
para$dry.totcnsmp.T35[para$dry.totcnsmp.T35 < 0] <- 1
para$dry.totcnsmp.T36[para$dry.totcnsmp.T36 < 0] <- 1
para$dry.totcnsmp.T37[para$dry.totcnsmp.T37 < 0] <- 1


```



Converting date into Julian day and calculating development time variables:

```{r}
#Parsing dates



para$date.hatch<-parse_date_time(para$date.hatch,orders='mdy')
para$date.died<-parse_date_time(para$date.died,orders='mdy')
para$date.3<-parse_date_time(para$date.3,orders='mdy')
para$date.ovp<-parse_date_time(para$date.ovp,orders='mdy')
para$date.4<-parse_date_time(para$date.4,orders='mdy')
para$date.5<-parse_date_time(para$date.5,orders='mdy')
para$date.wander<-parse_date_time(para$date.wander,orders='mdy')
para$date.em<-parse_date_time(para$date.em,orders='mdy')
para$date.T1<-parse_date_time(para$date.T1,orders='mdy')
para$date.T2<-parse_date_time(para$date.T2,orders='mdy')
para$date.T3<-parse_date_time(para$date.T3,orders='mdy')
para$date.T4<-parse_date_time(para$date.T4,orders='mdy')
para$date.T5<-parse_date_time(para$date.T5,orders='mdy')
para$date.T6<-parse_date_time(para$date.T6,orders='mdy')
para$date.T7<-parse_date_time(para$date.T7,orders='mdy')
para$date.T8<-parse_date_time(para$date.T8,orders='mdy')
para$date.T9<-parse_date_time(para$date.T9,orders='mdy')
para$date.T10<-parse_date_time(para$date.T10,orders='mdy')
para$date.T11<-parse_date_time(para$date.T11,orders='mdy')
para$date.T12<-parse_date_time(para$date.T12,orders='mdy')
para$date.T13<-parse_date_time(para$date.T13,orders='mdy')
para$date.T14<-parse_date_time(para$date.T14,orders='mdy')
para$date.T15<-parse_date_time(para$date.T15,orders='mdy')
para$date.T16<-parse_date_time(para$date.T16,orders='mdy')
para$date.T17<-parse_date_time(para$date.T17,orders='mdy')
para$date.T18<-parse_date_time(para$date.T18,orders='mdy')
para$date.T19<-parse_date_time(para$date.T19,orders='mdy')
para$date.T20<-parse_date_time(para$date.T20,orders='mdy')
para$date.T21<-parse_date_time(para$date.T21,orders='mdy')
para$date.T22<-parse_date_time(para$date.T22,orders='mdy')
para$date.T23<-parse_date_time(para$date.T23,orders='mdy')
para$date.T24<-parse_date_time(para$date.T24,orders='mdy')
para$date.T25<-parse_date_time(para$date.T25,orders='mdy')
para$date.T26<-parse_date_time(para$date.T26,orders='mdy')
para$date.T27<-parse_date_time(para$date.T27,orders='mdy')
para$date.T28<-parse_date_time(para$date.T28,orders='mdy')
para$date.T29<-parse_date_time(para$date.T29,orders='mdy')
para$date.T30<-parse_date_time(para$date.T30,orders='mdy')
para$date.T31<-parse_date_time(para$date.T31,orders='mdy')
para$date.T32<-parse_date_time(para$date.T32,orders='mdy')
para$date.T33<-parse_date_time(para$date.T33,orders='mdy')
para$date.T34<-parse_date_time(para$date.T34,orders='mdy')
para$date.T35<-parse_date_time(para$date.T35,orders='mdy')
para$date.T36<-parse_date_time(para$date.T36,orders='mdy')
para$date.T37<-parse_date_time(para$date.T37,orders='mdy')
para$date.T38<-parse_date_time(para$date.T38,orders='mdy')


#Converting to Julian days


para$date.ovp<-format(para$date.ovp, "%j")
para$date.ovp<-as.numeric(para$date.ovp)

para$date.em<-format(para$date.em, "%j")
para$date.em<-as.numeric(para$date.em)

para$date.hatch<-format(para$date.hatch,"%j")
para$date.hatch<-as.numeric(para$date.hatch)

para$date.died<-format(para$date.died,"%j")
para$date.died<-as.numeric(para$date.died)

para$date.3<-format(para$date.3,"%j")
para$date.3<-as.numeric(para$date.3)

para$date.4<-format(para$date.4,"%j")
para$date.4<-as.numeric(para$date.4)

para$date.5<-format(para$date.5,"%j")
para$date.5<-as.numeric(para$date.5)

para$date.wander<-format(para$date.wander,"%j")
para$date.wander<-as.numeric(para$date.wander)

para$date.T1<-format(para$date.T1,"%j")
para$date.T1<-as.numeric(para$date.T1)

para$date.T2<-format(para$date.T2,"%j")
para$date.T2<-as.numeric(para$date.T2)

para$date.T3<-format(para$date.T3,"%j")
para$date.T3<-as.numeric(para$date.T3)

para$date.T4<-format(para$date.T4,"%j")
para$date.T4<-as.numeric(para$date.T4)

para$date.T5<-format(para$date.T5,"%j")
para$date.T5<-as.numeric(para$date.T5)

para$date.T6<-format(para$date.T6,"%j")
para$date.T6<-as.numeric(para$date.T6)

para$date.T7<-format(para$date.T7,"%j")
para$date.T7<-as.numeric(para$date.T7)

para$date.T8<-format(para$date.T8,"%j")
para$date.T8<-as.numeric(para$date.T8)

para$date.T9<-format(para$date.T9,"%j")
para$date.T9<-as.numeric(para$date.T9)

para$date.T10<-format(para$date.T10,"%j")
para$date.T10<-as.numeric(para$date.T10)

para$date.T11<-format(para$date.T11,"%j")
para$date.T11<-as.numeric(para$date.T11)

para$date.T12<-format(para$date.T12,"%j")
para$date.T12<-as.numeric(para$date.T12)

para$date.T13<-format(para$date.T13,"%j")
para$date.T13<-as.numeric(para$date.T13)

para$date.T14<-format(para$date.T14,"%j")
para$date.T14<-as.numeric(para$date.T14)

para$date.T15<-format(para$date.T15,"%j")
para$date.T15<-as.numeric(para$date.T15)

para$date.T16<-format(para$date.T16,"%j")
para$date.T16<-as.numeric(para$date.T16)

para$date.T17<-format(para$date.T17,"%j")
para$date.T17<-as.numeric(para$date.T17)

para$date.T18<-format(para$date.T18,"%j")
para$date.T18<-as.numeric(para$date.T18)

para$date.T19<-format(para$date.T19,"%j")
para$date.T19<-as.numeric(para$date.T19)

para$date.T20<-format(para$date.T20,"%j")
para$date.T20<-as.numeric(para$date.T20)

para$date.T21<-format(para$date.T21,"%j")
para$date.T21<-as.numeric(para$date.T21)

para$date.T22<-format(para$date.T22,"%j")
para$date.T22<-as.numeric(para$date.T22)

para$date.T23<-format(para$date.T23,"%j")
para$date.T23<-as.numeric(para$date.T23)

para$date.T24<-format(para$date.T24,"%j")
para$date.T24<-as.numeric(para$date.T24)

para$date.T25<-format(para$date.T25,"%j")
para$date.T25<-as.numeric(para$date.T25)

para$date.T26<-format(para$date.T26,"%j")
para$date.T26<-as.numeric(para$date.T26)

para$date.T27<-format(para$date.T27,"%j")
para$date.T27<-as.numeric(para$date.T27)

para$date.T28<-format(para$date.T28,"%j")
para$date.T28<-as.numeric(para$date.T28)

para$date.T29<-format(para$date.T29,"%j")
para$date.T29<-as.numeric(para$date.T29)

para$date.T30<-format(para$date.T30,"%j")
para$date.T30<-as.numeric(para$date.T30)

para$date.T31<-format(para$date.T31,"%j")
para$date.T31<-as.numeric(para$date.T31)

para$date.T32<-format(para$date.T32,"%j")
para$date.T32<-as.numeric(para$date.T32)

para$date.T33<-format(para$date.T33,"%j")
para$date.T33<-as.numeric(para$date.T33)

para$date.T34<-format(para$date.T34,"%j")
para$date.T34<-as.numeric(para$date.T34)

para$date.T35<-format(para$date.T35,"%j")
para$date.T35<-as.numeric(para$date.T35)

para$date.T36<-format(para$date.T36,"%j")
para$date.T36<-as.numeric(para$date.T36)

para$date.T37<-format(para$date.T37,"%j")
para$date.T37<-as.numeric(para$date.T37)

para$date.T38<-format(para$date.T38,"%j")
para$date.T38<-as.numeric(para$date.T38)



para$int.wasp.dev<-para$date.em-para$date.ovp

para$timeto3<-para$date.3-para$date.hatch

para$int.3<-para$date.4<-para$date.3

para$timeto4<-para$timeto3+para$int.3

para$int.4<-para$date.5-para$date.4

para$timeto5<-para$timeto4+para$int.4

para$int.5.p<-para$date.em-para$date.5
  
para$int.5.c<-para$date.wander-para$date.5

para$int.5.c[is.na(para$int.5.c)] <- 0
para$int.5.p[is.na(para$int.5.p)] <- 0

para$int.5<-para$int.5.c+para$int.5.p

para$int.5[para$int.5==0]<-NA


para$ttwand<-para$date.wander-para$date.hatch


para$stsp.llmort<-para$num.fail.spin/para$num.em


```





Write data frame with consumption data into a csv file (will also manually add the new columns to the CPT full excel file)

```{r}
write.csv(para,"cpt gr wide.csv", row.names = FALSE, quote = FALSE)
```



NOTE: combined this and other cleaning scripts into one master cleaning script! Will keep this rmd as an explanation of the processes, but will use other script for actual data cleaning. 

NOTE: When making adjustments, from now on use raw cpt.csv--cpt gr wide now has julian dates and things that can't be converted back to raw format easily.
