---
title: "K/W lab meeting 3-17"
author: "Liz"
date: "May 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(gridExtra)
library(Rmisc)
library(readr)
library(nlme)

cpt <- read_csv("~/Manduca expts/Summer+Fall 2016/CxPxT/data files/cpt.gr.long.csv", 
    col_types = cols(temp = col_factor(levels = c("20", 
        "25", "30"))))
View(cpt)

wide <- read_csv("~/Manduca expts/Summer+Fall 2016/CxPxT/data files/cpt gr wide.csv", 
    col_types = cols(temp = col_factor(levels = c("20", 
        "25", "30")), treatment = col_factor(levels = c("control", 
        "para"))))
View(wide)


#Dividing by temperature

cpt.30<-subset(cpt,temp=="30")
View(cpt.30)

cpt.25<-subset(cpt,temp=="25")

cpt.20<-subset(cpt,temp=="20")


avg <- read_csv("~/Manduca expts/Summer+Fall 2016/CxPxT/data files/cpt avg data.csv")
View(avg)
avg$temp<-as.factor(avg$temp)
avg$treat<-as.factor(avg$treat)


avg.long <- read_csv("~/Manduca expts/Summer+Fall 2016/CxPxT/data files/cpt avg long.csv")
View(avg.long)
avg.long$temp<-as.factor(avg.long$temp)
avg.long$treat<-as.factor(avg.long$treat)



```



###Plot of actual age and mass, all temps and individuals

```{r,echo=FALSE,warning=FALSE}

k<- ggplot(cpt, aes(age, log(mass), group=interaction(bug.id, temp), color=factor(temp)))
k+geom_line(aes(linetype=treatment)
)+theme_bw(
)+theme(plot.background = element_blank(),panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.border= element_blank(), 
        axis.line.x = element_line(color="black", size = 1.5),
        axis.line.y = element_line(color="black", size = 1.5),text = element_text(size = 15)
)+scale_color_manual(values=c( "lightblue","darkblue", "magenta"),
                     guide=guide_legend(title="temperature"))

```

<br>
<br>
<br>


###Plot of actual age and consumption, all temps and individuals

```{r,warning=FALSE,echo=FALSE}

k<- ggplot(cpt, aes(age, log(tot.cnsmp), group=interaction(bug.id, temp), color=factor(temp)))
k+geom_line(aes(linetype=treatment)
)+theme_bw(
)+theme(plot.background = element_blank(),panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), panel.border= element_blank(), 
        axis.line.x = element_line(color="black", size = 1.5),
        axis.line.y = element_line(color="black", size = 1.5),text = element_text(size = 15)
)+scale_color_manual(values=c( "lightblue","darkblue", "magenta"),
                     guide=guide_legend(title="temperature"))

```

<br>
<br>
<br>



###Preliminary mixed effect models looking at mass, age, temperature and treatment

ID as a random intercept

```{r}

massmm.mod<-lme(log(mass)~temp*treatment*age,random=~1|bug.id,data=cpt,method="ML",na.action = na.omit)

anova(massmm.mod)

```

<br>
<br>
<br>

###Average mass by average age

```{r,echo=FALSE}

theme_set(theme_classic(base_size=14))

#Facet by treatment
avg.plot <- ggplot(avg.long, aes(x=age,y=log.mass,colour=temp,group=temp))
avg.plot<-avg.plot+geom_point()+geom_errorbar(aes(ymin=log.mass-log.mass.se,ymax=log.mass+log.mass.se),width=.2
)+geom_errorbarh(aes(xmin=age-age.se,xmax=age+age.se)
)+geom_line(aes(linetype=temp,color=temp),size=1
)+scale_color_manual(values=c("#56B4E9","#000000","#E69F00"), 
                     name="Temperature [C]",breaks=c("20","25","30"),
                     labels=c("20","25","30"),guide=guide_legend(keywidth=1.8)
)+scale_linetype_discrete(name="Temperature [C]",breaks=c("20","25","30"),
                          labels=c("20","25","30")
)+labs(title="Avg log(mass) X avg age",x="avg age [hour]",
       y="Avg log(mass) [mg]"
)+theme(axis.title.y=element_text(vjust=1.5),
        axis.title.x=element_text(vjust=-.35),
        plot.title=element_text(face="bold",vjust=1)
)+facet_wrap(~treat)



avg.plot

```

<br>
<br>
<br>

###Preliminary mixed effect models looking at consumption, age, temperature and treatment

ID as a random intercept
<br>
<br>


```{r}
cnsmp.mod<-lme(log(tot.cnsmp)~treatment*temp*age,random=~1|bug.id,data=cpt,method="ML",na.action = na.omit)
anova(cnsmp.mod)
```
<br>
<br>
<br>

###Average consumption by average age

```{r,echo=FALSE}
#Facet by treatment
avg.plot <- ggplot(avg.long, aes(x=age,y=log.cnsmp,colour=temp,group=temp))
avg.plot<-avg.plot+geom_point()+geom_errorbar(aes(ymin=log.cnsmp-log.cnsmp.se,ymax=log.cnsmp+log.cnsmp.se),width=.2
)+geom_errorbarh(aes(xmin=age-age.se,xmax=age+age.se)
)+geom_line(aes(linetype=temp,color=temp),size=1
)+scale_color_manual(values=c("#56B4E9","#000000","#E69F00"), 
                     name="Temperature [C]",breaks=c("20","25","30"),
                     labels=c("20","25","30"),guide=guide_legend(keywidth=1.8)
)+scale_linetype_discrete(name="Temperature [C]",breaks=c("20","25","30"),
                          labels=c("20","25","30")
)+scale_y_continuous(limits=c(0,9),breaks = c(4,6,8)
)+labs(title="Avg log(cnsmp) X avg age",x="avg age [hour]",
       y="Avg log(cnsmp) [mg]"
)+theme(axis.title.y=element_text(vjust=1.5),
        axis.title.x=element_text(vjust=-.35),
        plot.title=element_text(face="bold",vjust=1)
)+facet_wrap(~treat)

avg.plot


```

<br>
<br>
<br>

###Caterpillar development time to the 4th instar:

####Linear model:


```{r}
tt4.mod2<-lm(timeto4~temp*treatment+num.em,data=wide,na.action = na.omit)
anova(tt4.mod2)

```


```{r}
summary(tt4.mod2)

```

<br>
<br>
<br>

####Average development time to the 4th instar by temperature

```{r,echo=FALSE}
tt4.sum<-summarySE(wide, measurevar="timeto4", 
                     groupvars=c("temp","treatment"),na.rm=TRUE)
tt4.sum


tt4plot <- ggplot(tt4.sum, aes(x=temp,y=timeto4,group=treatment,color=treatment))
tt4plot<-tt4plot+geom_point()+geom_errorbar(aes(ymin=timeto4-se,ymax=timeto4+se),width=.2
)+geom_line(aes(linetype=treatment,color=treatment)
)+scale_color_manual(values=c("black","red"), 
                     name="Treatment",breaks=c("control","wide"),
                     labels=c("Control","widesitized"),guide=guide_legend(keywidth=1.8)
)+scale_linetype_discrete(name="Treatment",breaks=c("control","wide"),
                          labels=c("Control","widesitized")
)+labs(title="Dev time to 4th instar",x="Temperature [C]",
       y="Caterpillar dev [days]"
)+theme(axis.title.y=element_text(vjust=1.5),
        axis.title.x=element_text(vjust=-.35),
        plot.title=element_text(face="bold",vjust=1))

tt4plot


```

<br>
<br>
<br>



### time to 5 model

```{r}
tt5.mod2<-lm(timeto5~temp*treatment+num.em,data=wide,na.action = na.omit)
anova(tt5.mod2)


```

```{r}
summary(tt5.mod2)
```

<br>
<br>
<br>


###Plot of caterpillar development time to the 5th instar

```{r,echo=FALSE}
tt5.sum<-summarySE(wide, measurevar="timeto5", 
                   groupvars=c("temp","treatment"),na.rm=TRUE)
tt5.sum


tt5plot <- ggplot(tt5.sum, aes(x=temp,y=timeto5,group=treatment,color=treatment))
tt5plot<-tt5plot+geom_point()+geom_errorbar(aes(ymin=timeto5-se,ymax=timeto5+se),width=.2
)+geom_line(aes(linetype=treatment,color=treatment)
)+scale_color_manual(values=c("black","red"), 
                     name="Treatment",breaks=c("control","wide"),
                     labels=c("Control","widesitized"),guide=guide_legend(keywidth=1.8)
)+scale_linetype_discrete(name="Treatment",breaks=c("control","wide"),
                          labels=c("Control","widesitized")
)+labs(title="Dev time to 5th instar",x="Temperature [C]",
       y="Caterpillar dev [days]"
)+theme(axis.title.y=element_text(vjust=1.5),
        axis.title.x=element_text(vjust=-.35),
        plot.title=element_text(face="bold",vjust=1))

tt5plot

```

<br>
<br>
<br>



###Development time to wandering in control caterpillars

```{r,include=FALSE}
control<-subset(wide,treatment=="control")
```



```{r}
ttw.mod<-lm(ttwand~temp,data=control,na.action = na.omit)
anova(ttw.mod)
```

```{r,echo=FALSE}
ttw.plot<-ggplot(control,aes(x=temp,y=ttwand))
ttw.plot<-ttw.plot+geom_boxplot()
ttw.plot
```

<br>
<br>
<br>



###Time to wasp emergence


```{r}
wm.mod<-lm(int.wasp.dev~temp*num.em,data=wide,na.action = na.omit)
anova(wm.mod)
```

```{r}

summary(wm.mod)
```


<br>
<br>
<br>


###Average wasp development time

```{r,echo=FALSE}
wm.sum<-summarySE(wide, measurevar="int.wasp.dev", 
                  groupvars=c("temp"),na.rm=TRUE)
wm.sum

wmplot <- ggplot(wm.sum, aes(x=temp,y=int.wasp.dev,group=1))
wmplot<-wmplot+geom_point()+geom_errorbar(aes(ymin=int.wasp.dev-se,ymax=int.wasp.dev+se),width=.2
)+geom_line(
)+labs(title="Wasp dev",x="Temperature [C]",
       y="wasp dev [days]"
)+theme(axis.title.y=element_text(vjust=1.5),
        axis.title.x=element_text(vjust=-.35),
        plot.title=element_text(face="bold",vjust=1))

wmplot
```

<br>
<br>
<br>

###Wasp development by number emerged

```{r,echo=FALSE}
wmplot<-ggplot(wide,aes(x=num.em,y=int.wasp.dev,group=temp,color=temp))
wmplot<-wmplot+geom_point()+scale_color_manual(values=c("#56B4E9","#000000","#E69F00"), 
                                               name="Temperature [C]",breaks=c("20","25","30"),
                                               labels=c("20","25","30"),guide=guide_legend(keywidth=1.8)
)+geom_jitter()+geom_smooth(method=lm,se=FALSE)
wmplot

```

<br>
<br>
<br>



###Wasp late larval mortality

```{r,include=FALSE}
paratrt<-subset(wide,treatment=="para")
```



```{r}
stsllmort.mod<-lm(stsp.llmort~temp*num.em,data=paratrt,na.action = na.omit)
anova(stsllmort.mod)
```

```{r}
summary(stsllmort.mod)
```


```{r,echo=FALSE}
ssllm.plot<-ggplot(paratrt,aes(x=temp,y=stsp.llmort))
ssllm.plot<-ssllm.plot+geom_boxplot()
ssllm.plot
```


























