---
title: "Extracting values from matching rows and columns (consumption)"
author: "Liz"
date: "May 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Extracting consumption for important time points (4th instar, 5th instar, wandering, emergence)



Total consumption at 4th instar

Subset to only identifying columns (num, bug.id), timepoint at 4th instar, and consumption columns
```{r}

keepvars<-c("num","bug.id","timepoint.4","dry.totcnsmp.T1","dry.totcnsmp.T2","dry.totcnsmp.T3","dry.totcnsmp.T4","dry.totcnsmp.T5","dry.totcnsmp.T6","dry.totcnsmp.T7","dry.totcnsmp.T8","dry.totcnsmp.T9","dry.totcnsmp.T10","dry.totcnsmp.T11","dry.totcnsmp.T12","dry.totcnsmp.T13","dry.totcnsmp.T14","dry.totcnsmp.T15","dry.totcnsmp.T16","dry.totcnsmp.T17","dry.totcnsmp.T18","dry.totcnsmp.T19","dry.totcnsmp.T20","dry.totcnsmp.T21","dry.totcnsmp.T22","dry.totcnsmp.T23","dry.totcnsmp.T24","dry.totcnsmp.T25","dry.totcnsmp.T26","dry.totcnsmp.T27","dry.totcnsmp.T28","dry.totcnsmp.T29","dry.totcnsmp.T30","dry.totcnsmp.T31","dry.totcnsmp.T32","dry.totcnsmp.T33","dry.totcnsmp.T34","dry.totcnsmp.T35","dry.totcnsmp.T36","dry.totcnsmp.T37")
sub<-para[keepvars]
View(sub)
sub<-as.data.frame(sub)
```



Rename (using dplyr) columns so that they match the timepoint designations in timepoint.4 column (T1, T2, etc)

```{r}
sub<-rename(sub,T1=dry.totcnsmp.T1,T2=dry.totcnsmp.T2,T3=dry.totcnsmp.T3,T4=dry.totcnsmp.T4,T5=dry.totcnsmp.T5,T6=dry.totcnsmp.T6,T7=dry.totcnsmp.T7,T8=dry.totcnsmp.T8,T9=dry.totcnsmp.T9,T10=dry.totcnsmp.T10,T11=dry.totcnsmp.T11,T12=dry.totcnsmp.T12,T13=dry.totcnsmp.T13,T14=dry.totcnsmp.T14,T15=dry.totcnsmp.T15,T16=dry.totcnsmp.T16,T17=dry.totcnsmp.T17,T18=dry.totcnsmp.T18,T19=dry.totcnsmp.T19,T20=dry.totcnsmp.T20,T21=dry.totcnsmp.T21,T22=dry.totcnsmp.T22,T23=dry.totcnsmp.T23,T24=dry.totcnsmp.T24,T25=dry.totcnsmp.T25,T26=dry.totcnsmp.T26,T27=dry.totcnsmp.T27,T28=dry.totcnsmp.T28,T29=dry.totcnsmp.T29,T30=dry.totcnsmp.T30,T31=dry.totcnsmp.T31,T32=dry.totcnsmp.T32,T33=dry.totcnsmp.T33,T34=dry.totcnsmp.T34,T35=dry.totcnsmp.T35,T36=dry.totcnsmp.T36,T37=dry.totcnsmp.T37)

```


Create a column index, match it with a row index, extract the matching data (timepoint.4 row value with row value of corresponding consumption column), and put that data into a new column in the data set  ([] tells it to look in all columns)

```{r}
sub$cnsmp.4 <- sub[][cbind(1:nrow(sub), match(sub$timepoint.4, colnames(sub)))]

```

Subset to only identifiers and new consumption column

```{r}
keepvars2<-c("num","bug.id","cnsmp.4")
sub2<-sub[keepvars2]
sub2<-as.data.frame(sub2)

```


Merge with full data set by identifiers. Arrange by number, check the rows match in new column value

```{r}
para<-merge(para,sub2,by=c("num","bug.id"))

para<-arrange(para, num)

para[c("num", "cnsmp.4")]
sub2[c("num", "cnsmp.4")]

```




Total consumption at 5th instar

Subset to only identifying columns (num, bug.id), timepoint at 5th instar, and consumption columns

```{r}
keepvars3<-c("num","bug.id","timepoint.5","dry.totcnsmp.T1","dry.totcnsmp.T2","dry.totcnsmp.T3","dry.totcnsmp.T4","dry.totcnsmp.T5","dry.totcnsmp.T6","dry.totcnsmp.T7","dry.totcnsmp.T8","dry.totcnsmp.T9","dry.totcnsmp.T10","dry.totcnsmp.T11","dry.totcnsmp.T12","dry.totcnsmp.T13","dry.totcnsmp.T14","dry.totcnsmp.T15","dry.totcnsmp.T16","dry.totcnsmp.T17","dry.totcnsmp.T18","dry.totcnsmp.T19","dry.totcnsmp.T20","dry.totcnsmp.T21","dry.totcnsmp.T22","dry.totcnsmp.T23","dry.totcnsmp.T24","dry.totcnsmp.T25","dry.totcnsmp.T26","dry.totcnsmp.T27","dry.totcnsmp.T28","dry.totcnsmp.T29","dry.totcnsmp.T30","dry.totcnsmp.T31","dry.totcnsmp.T32","dry.totcnsmp.T33","dry.totcnsmp.T34","dry.totcnsmp.T35","dry.totcnsmp.T36","dry.totcnsmp.T37")
sub3<-para[keepvars3]
sub3<-as.data.frame(sub3)
```


Rename (using dplyr) columns so that they match the timepoint designations in timepoint.5 column (T1, T2, etc)

```{r}
sub3<-rename(sub3, T1=dry.totcnsmp.T1,T2=dry.totcnsmp.T2,T3=dry.totcnsmp.T3,T4=dry.totcnsmp.T4,T5=dry.totcnsmp.T5,T6=dry.totcnsmp.T6,T7=dry.totcnsmp.T7,T8=dry.totcnsmp.T8,T9=dry.totcnsmp.T9,T10=dry.totcnsmp.T10,T11=dry.totcnsmp.T11,T12=dry.totcnsmp.T12,T13=dry.totcnsmp.T13,T14=dry.totcnsmp.T14,T15=dry.totcnsmp.T15,T16=dry.totcnsmp.T16,T17=dry.totcnsmp.T17,T18=dry.totcnsmp.T18,T19=dry.totcnsmp.T19,T20=dry.totcnsmp.T20,T21=dry.totcnsmp.T21,T22=dry.totcnsmp.T22,T23=dry.totcnsmp.T23,T24=dry.totcnsmp.T24,T25=dry.totcnsmp.T25,T26=dry.totcnsmp.T26,T27=dry.totcnsmp.T27,T28=dry.totcnsmp.T28,T29=dry.totcnsmp.T29,T30=dry.totcnsmp.T30,T31=dry.totcnsmp.T31,T32=dry.totcnsmp.T32,T33=dry.totcnsmp.T33,T34=dry.totcnsmp.T34,T35=dry.totcnsmp.T35,T36=dry.totcnsmp.T36,T37=dry.totcnsmp.T37)

```


Create a column index, match it with a row index, extract the matching data (timepoint.5 row value with row value of corresponding consumption column), and put that data into a new column in the data set  ([] tells it to look in all columns)


```{r}
sub3$cnsmp.5 <- sub3[][cbind(1:nrow(sub3), match(sub3$timepoint.5, colnames(sub3)))]
```



Subset to only identifiers and new consumption column

```{r}
keepvars3.5<-c("num","bug.id","cnsmp.5")
sub3.5<-sub3[keepvars3.5]
sub3.5<-as.data.frame(sub3.5)

```


Merge with full data set by identifiers. Arrange by number, check the rows match in new column value


```{r}
para<-merge(para,sub3.5,by=c("num","bug.id"))

para<-arrange(para, num)

para[c("num", "cnsmp.5")]
sub3[c("num", "cnsmp.5")]

```


Total consumption at wandering

Subset to only identifying columns (num, bug.id), timepoint at wandering, and consumption columns

```{r}
keepvars4<-c("num","bug.id","timepoint.wander","dry.totcnsmp.T1","dry.totcnsmp.T2","dry.totcnsmp.T3","dry.totcnsmp.T4","dry.totcnsmp.T5","dry.totcnsmp.T6","dry.totcnsmp.T7","dry.totcnsmp.T8","dry.totcnsmp.T9","dry.totcnsmp.T10","dry.totcnsmp.T11","dry.totcnsmp.T12","dry.totcnsmp.T13","dry.totcnsmp.T14","dry.totcnsmp.T15","dry.totcnsmp.T16","dry.totcnsmp.T17","dry.totcnsmp.T18","dry.totcnsmp.T19","dry.totcnsmp.T20","dry.totcnsmp.T21","dry.totcnsmp.T22","dry.totcnsmp.T23","dry.totcnsmp.T24","dry.totcnsmp.T25","dry.totcnsmp.T26","dry.totcnsmp.T27","dry.totcnsmp.T28","dry.totcnsmp.T29","dry.totcnsmp.T30","dry.totcnsmp.T31","dry.totcnsmp.T32","dry.totcnsmp.T33","dry.totcnsmp.T34","dry.totcnsmp.T35","dry.totcnsmp.T36","dry.totcnsmp.T37")
sub4<-para[keepvars4]
sub4<-as.data.frame(sub4)

```


Rename (using dplyr) columns so that they match the timepoint designations in timepoint.wander column (T1, T2, etc)


```{r}
sub4<-rename(sub4, T1=dry.totcnsmp.T1,T2=dry.totcnsmp.T2,T3=dry.totcnsmp.T3,T4=dry.totcnsmp.T4,T5=dry.totcnsmp.T5,T6=dry.totcnsmp.T6,T7=dry.totcnsmp.T7,T8=dry.totcnsmp.T8,T9=dry.totcnsmp.T9,T10=dry.totcnsmp.T10,T11=dry.totcnsmp.T11,T12=dry.totcnsmp.T12,T13=dry.totcnsmp.T13,T14=dry.totcnsmp.T14,T15=dry.totcnsmp.T15,T16=dry.totcnsmp.T16,T17=dry.totcnsmp.T17,T18=dry.totcnsmp.T18,T19=dry.totcnsmp.T19,T20=dry.totcnsmp.T20,T21=dry.totcnsmp.T21,T22=dry.totcnsmp.T22,T23=dry.totcnsmp.T23,T24=dry.totcnsmp.T24,T25=dry.totcnsmp.T25,T26=dry.totcnsmp.T26,T27=dry.totcnsmp.T27,T28=dry.totcnsmp.T28,T29=dry.totcnsmp.T29,T30=dry.totcnsmp.T30,T31=dry.totcnsmp.T31,T32=dry.totcnsmp.T32,T33=dry.totcnsmp.T33,T34=dry.totcnsmp.T34,T35=dry.totcnsmp.T35,T36=dry.totcnsmp.T36,T37=dry.totcnsmp.T37)

```


Since food wasn't weighed at the timepoint where they wandered, have to subtract 1 to get the last total consumption measured. Use gsub to take out the T from the timepoint.wander column, change to numeric to minus 1, then use gsub to put the T back in the column. ^ is regex expression to for space in front of something

```{r}
sub4$timepoint.wander<-gsub("T", "",sub4$timepoint.wander)
sub4$timepoint.wander<-as.numeric(sub4$timepoint.wander)-1
sub4$timepoint.wander<-gsub("^", "T",sub4$timepoint.wander)
```


Create a column index, match it with a row index, extract the matching data (timepoint.5 row value with row value of corresponding consumption column), and put that data into a new column in the data set  ([] tells it to look in all columns)


```{r}
sub4$cnsmp.wan <- sub4[][cbind(1:nrow(sub4), match(sub4$timepoint.wander, colnames(sub4)))]

```


Subset to only identifiers and new consumption column

```{r}

keepvars4.5<-c("num","bug.id","cnsmp.wan")
sub4.5<-sub4[keepvars4.5]
sub4.5<-as.data.frame(sub4.5)

```


Merge with full data set by identifiers. Arrange by number, check the rows match in new column value

```{r}
para<-merge(para,sub4.5,by=c("num","bug.id"))

para<-arrange(para, num)

para[c("num", "cnsmp.wan")]
sub4[c("num", "cnsmp.wan")]
```




Total consumption at wasp emergence

Subset to only identifying columns (num, bug.id), timepoint at emergence, and consumption columns

```{r}

keepvars5<-c("num","bug.id","timepoint.em","dry.totcnsmp.T1","dry.totcnsmp.T2","dry.totcnsmp.T3","dry.totcnsmp.T4","dry.totcnsmp.T5","dry.totcnsmp.T6","dry.totcnsmp.T7","dry.totcnsmp.T8","dry.totcnsmp.T9","dry.totcnsmp.T10","dry.totcnsmp.T11","dry.totcnsmp.T12","dry.totcnsmp.T13","dry.totcnsmp.T14","dry.totcnsmp.T15","dry.totcnsmp.T16","dry.totcnsmp.T17","dry.totcnsmp.T18","dry.totcnsmp.T19","dry.totcnsmp.T20","dry.totcnsmp.T21","dry.totcnsmp.T22","dry.totcnsmp.T23","dry.totcnsmp.T24","dry.totcnsmp.T25","dry.totcnsmp.T26","dry.totcnsmp.T27","dry.totcnsmp.T28","dry.totcnsmp.T29","dry.totcnsmp.T30","dry.totcnsmp.T31","dry.totcnsmp.T32","dry.totcnsmp.T33","dry.totcnsmp.T34","dry.totcnsmp.T35","dry.totcnsmp.T36","dry.totcnsmp.T37")
sub5<-para[keepvars5]
sub5<-as.data.frame(sub5)

```



Rename (using dplyr) columns so that they match the timepoint designations in timepoint.wander column (T1, T2, etc)

```{r}
sub5<-rename(sub5, T1=dry.totcnsmp.T1,T2=dry.totcnsmp.T2,T3=dry.totcnsmp.T3,T4=dry.totcnsmp.T4,T5=dry.totcnsmp.T5,T6=dry.totcnsmp.T6,T7=dry.totcnsmp.T7,T8=dry.totcnsmp.T8,T9=dry.totcnsmp.T9,T10=dry.totcnsmp.T10,T11=dry.totcnsmp.T11,T12=dry.totcnsmp.T12,T13=dry.totcnsmp.T13,T14=dry.totcnsmp.T14,T15=dry.totcnsmp.T15,T16=dry.totcnsmp.T16,T17=dry.totcnsmp.T17,T18=dry.totcnsmp.T18,T19=dry.totcnsmp.T19,T20=dry.totcnsmp.T20,T21=dry.totcnsmp.T21,T22=dry.totcnsmp.T22,T23=dry.totcnsmp.T23,T24=dry.totcnsmp.T24,T25=dry.totcnsmp.T25,T26=dry.totcnsmp.T26,T27=dry.totcnsmp.T27,T28=dry.totcnsmp.T28,T29=dry.totcnsmp.T29,T30=dry.totcnsmp.T30,T31=dry.totcnsmp.T31,T32=dry.totcnsmp.T32,T33=dry.totcnsmp.T33,T34=dry.totcnsmp.T34,T35=dry.totcnsmp.T35,T36=dry.totcnsmp.T36,T37=dry.totcnsmp.T37)

```


Since food wasn't weighed at the timepoint where the wasps emerged, have to subtract 1 to get the last total consumption measured. Use gsub to take out the T from the timepoint.em column, change to numeric to minus 1, then use gsub to put the T back in the column. ^ is regex expression to for space in front of something

```{r}
sub5$timepoint.em<-gsub("T", "",sub5$timepoint.em)
sub5$timepoint.em<-as.numeric(sub5$timepoint.em)-1
sub5$timepoint.em<-gsub("^", "T",sub5$timepoint.em)

```



Create a column index, match it with a row index, extract the matching data (timepoint.5 row value with row value of corresponding consumption column), and put that data into a new column in the data set  ([] tells it to look in all columns)

```{r}
sub5$cnsmp.em <- sub5[][cbind(1:nrow(sub5), match(sub5$timepoint.em, colnames(sub5)))]

```



Subset to only identifiers and new consumption column

```{r}
keepvars5.5<-c("num","bug.id","cnsmp.em")
sub5.5<-sub5[keepvars5.5]
sub5.5<-as.data.frame(sub5.5)

```


Merge with full data set by identifiers. Arrange by number, check the rows match in new column value

```{r}
para<-merge(para,sub5.5,by=c("num","bug.id"))

para<-arrange(para, num)

para[c("num", "cnsmp.em")]
sub5[c("num", "cnsmp.em")]
```


